RegressionPvNP

### 2017-05-16 RegressionPvNP.Rmd
fixing loads of little mistakes that results in messing up the pipeline entirely.

sum.patient.data = summarise.patient.info(patient.info)
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now

@@

# Missing some of the samples as they aren't all sequenced yet
for (pt in names(patient.data)) {
  #print(pt)
  #print(patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)])
  patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
}
```

@@

tile.sample.segments<-function(tileFile, sample.data) {
  if (file.exists(tileFile)) {
    message(paste("Loading", tileFile))
    load(tileFile, verbose=T)
  } else {
    message("Tiling patient data...")
    grList = lapply(sample.data, function(df) {
      df$info = arrange(df$info, Endoscopy.Year, Pathology)

@@

mergedDf = tile.sample.segments(paste(as.character(tile.w),'_tiledpts.Rdata', sep=''), patient.data)

#validatioMergedDf = tile.sample.segments(paste(as.character(tile.w),'validation_tiledpts.Rdata', sep=''), validation.patient.data)

@@

dysplasia.df = t(mergedDf[,names(labels)])
```

@@

Example subset of the data matrix:
`r pander(dysplasia.df[1:10, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

@@

crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='performance') {
  if (nfolds > 5) minR = 0.1
  message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
  fit.e = list()
  tpts = create.patient.sets(pts, nfolds, splits, minR)
  cv.pred = (matrix(nrow=0, ncol=length(lambda)))
  cv.binomial.deviance = (matrix(nrow=0, ncol=length(lambda)))
  for (n in 1:nfolds) {  
    message(paste(n, "fold"))
    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(tpts))

@@

      # Binomial deviance, lifted from cv.lognet
      predmat = predict(fitCV, test, type='response')
      prob_min = 1e-05; prob_max = 1 - prob_min

      predmat = pmin(pmax(predmat, prob_min), prob_max)

      yT = as.factor(y[test.rows])
      dev = apply(predmat, 2, function(p)  -2*((yT==1)*log(p)+(yT==0)*log(1-p)) )

      deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)

@@

  # Not really sure what to do with the binomial deviance now...
  df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
                        'mean'= colMeans(cv.pred), 
                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
                        'sd'= apply(cv.pred, 2, sd), 
                        'lambda'= lambda,
                        'log.lambda' = log(lambda),
                        'mean.b.dev' = colMeans(cv.binomial.deviance),
                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
                        'lambda.min' = F, 'lambda.1se' = F, `lambda-1se` = F)
  #df$lambda.1se = FALSE
  #df$`lambda-1se` = FALSE

  if (select == 'performance') {
    df$lambda.min = df$mean == max(df$mean)
    df[df$log.lambda < subset(df, mean == max(mean))$log.lambda+sd(df$log.lambda),][1, 'lambda.1se'] = TRUE
    #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
  } else if (select == 'deviance') {
    df$lambda.min = df$mean.b.dev == min(df$mean.b.dev)
    df[df$mean.b.dev >= df[df$lambda.min == T, 'mean.b.dev'] + sd(df$mean.b.dev), ][1, 'lambda.1se'] = T
  }

@@

  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 
              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
}

@@

  if (nrow(subset(df, lambda.1se == T)) > 0) {
    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.min+1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3))

    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min+1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3)) 
  }

  if (nrow(subset(df, `lambda-1se` == T)) > 0) {
      gp = gp + geom_point(data=subset(df, `lambda-1se` == T), aes(y=mean, x=log.lambda, colour="lambda.min-1se"), size=2 ) +
      geom_vline(xintercept = subset(df, `lambda-1se` == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, `lambda-1se` == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, `lambda-1se` == T)$lambda.at,'mean'],3))

      gpD = gpD + geom_point(data=subset(df, `lambda-1se` == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min-1se"), size=2 ) +
      geom_vline(xintercept = subset(df, `lambda-1se` == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, `lambda-1se` == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, `lambda-1se` == T)$lambda.at,'mean'],3)) 
  }

@@

    coef.min = subset(df, lambda.min == T)$nzcoef
    coef.1se = subset(df, lambda.1se == T)$nzcoef
    coef.m.1se = subset(df, `lambda-1se` == T)$nzcoef

@@

    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, coef.m.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
    coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)

@@

```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
performance.plots = list(); coefs = list(); deviance.plots = list()
plots = list()
folds = 10; splits=5
alpha.values = c(0,0.5,0.7,0.8,0.9,1)

if (file.exists('/tmp/all.pt.alpha.Rdata')) {
  load('/tmp/all.pt.alpha.Rdata', verbose=F)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=1000, family='binomial') # all patients
    autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance')

@@

    #performance.plots[[as.character(a)]] = cv.patient$plot + ggtitle(paste('alpha=',a,sep=''))
    #deviance.plots[[as.character(a)]] = cv.patient$deviance.plot + ggtitle(paste('alpha=',a,sep=''))

    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
  }
  save(plots, coefs, file='/tmp/all.pt.alpha.Rdata')
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

@@

if (file.exists('/tmp/nohgd.Rdata')) {
  load('/tmp/nohgd.Rdata', verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)

  no.hgd.plots = list()
  coefs = list()
  for (a in alpha.values) {
    fit1 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
    autoplot(fit1, xvar='lambda', main='fit1, no HGD/IMC samples') + theme(legend.position='none') 
    lambda1 <- fit1$lambda

    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda1, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit1, select='deviance')

    #plots[[as.character(a)]] = cv.nohgd$plot + ggtitle(paste('alpha=',a,sep=''))

    no.hgd.plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fit1, cv.nohgd$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nohgd$non.zero.cf)
  }
  save(no.hgd.plots, coefs, file='/tmp/nohgd.Rdata')
}

do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))

@@

```{r noLGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}

if (file.exists('/tmp/nolgd.Rdata')) {
  load('/tmp/nolgd.Rdata', verbose=T)
} else {
  # No LGD
  samples = intersect(rownames(dysplasia.df), c(subset(info, Status == 'NP')$Samplename, subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD') & Status == 'P')$Samplename))

@@

  nolgd.plots = list(); coefs = list()
  for (a in alpha.values) {
    fit2 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
    autoplot(fit2, xvar='lambda', main='fit1, no HGD/IMC samples') + theme(legend.position='none') 
    lambda2 <- fit2$lambda

    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda2, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit2, select='deviance')

    #plots[[as.character(a)]] = cv.nolgd$plot + ggtitle(paste('alpha=',a,sep=''))

    nolgd.plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('CV Performance'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fit2, cv.nolgd$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)
  }
  save(nolgd.plots, coefs, file='/tmp/nolgd.Rdata')
}

do.call(grid.arrange, c(plots, top="No HGD/IMC/LGD samples", ncol=2))

@@

```{r, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
m = median(subset(sum.patient.data, Status == 'P')$total.samples)

nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$info$Samplename)))

# Few
samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples < m)$Patient], function(pt) pt$info$Samplename))))

few.plots = list()
for (a in alpha.values) {
  fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
  cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, minR=0.1, select='deviance')
  #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))

  few.plots[[as.character(a)]] = arrangeGrob(cv.back.few$plot+ggtitle('CV Performance'), cv.back.few$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
}
do.call(grid.arrange, c(few.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples < m))," Progressors with < median # samples (",m,")")))

# Many
samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Patient], function(pt) pt$info$Samplename))))

many.plots = list()
for (a in alpha.values) {
  fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients

  cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, select='deviance')
  #many.plots[[as.character(a)]] = cv.back.many$plot + ggtitle(paste('alpha=',a,sep=''))

  many.plots[[as.character(a)]] = arrangeGrob(cv.back.many$plot+ggtitle('CV Performance'), cv.back.many$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
}
do.call(grid.arrange, c(many.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m))," Progressors with >= median # samples (",m,")")))

@@

info = do.call(rbind, lapply(patient.data, function(df) df$info))
pg.samp = lapply(patient.data, function(pt) {
  info = pt$info
  info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
  info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
  info$Prediction = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

if (file.exists('/tmp/loo.Rdata')) {
  load('/tmp/loo.Rdata', verbose=T)
} else {

  performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
  # Remove each patient (LOO)
  for (pt in names(pg.samp)) {
    tmp.patient.data = patient.data[subset(sum.patient.data, Patient != pt)$Patient]
    samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))
    length(samples)

    # Fit generated on all samples, including HGD
    a = 1
    fitLOO <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
    cv = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitLOO$lambda, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO, select='deviance')
    #plots[[pt]] = cv$plot + ggtitle(pt)

    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('CV Performance'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)

@@

predicted.progression<-function(x) { if(is.na(x)) return(NA); as.integer(length(which(x == 1)) > 0)  }

@@

What about the non-progressors that had progressor predictions?
```{r}

@@

>>>>

sum.patient.data = summarise.patient.info(patient.info)

if (length(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)) > 0)
  warning(paste("Patients missing data:",
    paste(names(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)), collapse=', ')))

patient.data[which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)] = NULL
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now

# Missing some of the samples as they aren't all sequenced yet
for (pt in names(patient.data)) {
  #print(pt)
  #print(patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)])
  patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
}

@@

@@

tile.sample.segments<-function(tileFile, sample.data) {
  if (file.exists(tileFile)) {
    message(paste("Loading", tileFile))
    load(tileFile, verbose=T)
  } else {
    message("Tiling patient data...")
    grList = lapply(sample.data, function(df) {
      print(unique(df$info$Patient))
      df$info = arrange(df$info, Endoscopy.Year, Pathology)

@@

mergedDf = tile.sample.segments(paste(as.character(tile.w),'_tiledpts.Rdata', sep=''), patient.data)
dim(mergedDf)
#validatioMergedDf = tile.sample.segments(paste(as.character(tile.w),'validation_tiledpts.Rdata', sep=''), validation.patient.data)

@@

dysplasia.df = t(mergedDf[,names(labels)])
dim(dysplasia.df)
```

@@

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

@@

binomial.deviance<-function(pmat, y) {
  # Binomial deviance, lifted from cv.lognet
  prob_min = 1e-05; prob_max = 1 - prob_min
  pmat = pmin(pmax(pmat, prob_min), prob_max)
  dev = apply(pmat, 2, function(p)  -2*((y==1)*log(p)+(y==0)*log(1-p)) )
  return(dev)
}

crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='classification') {
  if (nfolds > 5) minR = 0.1
  message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
  fit.e = list()
  tpts = create.patient.sets(pts, nfolds, splits, minR)
  cv.pred = (matrix(nrow=0, ncol=length(lambda)))
  cv.binomial.deviance = (matrix(nrow=0, ncol=length(lambda)))
  for (n in 1:nfolds) {  
    message(paste(n, "fold"))
    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(tpts))

@@

      # Binomial deviance, lifted from cv.lognet
      dev = binomial.deviance(predict(fitCV, test, type='response'), as.factor(y[test.rows]))
      deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)

@@

  # Not really sure what to do with the binomial deviance now...
  df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
                        'mean'= colMeans(cv.pred), 
                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
                        'sd'= apply(cv.pred, 2, sd), 
                        'lambda'= lambda,
                        'log.lambda' = log(lambda),
                        'mean.b.dev' = colMeans(cv.binomial.deviance),
                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
                        'lambda.min' = F, 'lambda.1se' = F)

  if (select == 'classification') {
    df$lambda.min = df$mean == max(df$mean)
    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']

    arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1:10,]
    df[arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1, 'lambda.at'], 'lambda.1se'] = T
    #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
  } else if (select == 'deviance') {
    df$lambda.min = df$mean.b.dev == min(df$mean.b.dev)
    se1 = df[df$lambda.min == T, 'mean.b.dev'] - df[df$lambda.min == T, 'sme.b.dev']
    df[arrange(subset(df, mean.b.dev <= se1 & !lambda.min), -mean.b.dev)[1, 'lambda.at'], 'lambda.1se'] = T
    #df[df$mean.b.dev >= df[df$lambda.min == T, 'mean.b.dev'] + sd(df$mean.b.dev), ][1, 'lambda.1se'] = T
  }

@@

  return(list('max.cm'=df[df$lambda.1se ==T,'mean'], 
              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
}

@@

  if (nrow(subset(df, lambda.1se == T)) > 0) {
    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3))

    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3)) 
  }

@@

    coef.min = subset(df, lambda.min == T)$nzcoef
    coef.1se = subset(df, lambda.1se == T)$nzcoef

@@

    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
    coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)

@@

```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
coefs = list(); plots = list(); performance.at.1se = list()
folds = 10; splits = 5
alpha.values = c(0,0.5,0.7,0.8,0.9,1)
if (file.exists('/tmp/all.pt.alpha.Rdata')) {
  load('/tmp/all.pt.alpha.Rdata', verbose=F)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=1000, family='binomial') # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)

@@

    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('Classification'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
  }
  save(plots, coefs, performance.at.1se, file='/tmp/all.pt.alpha.Rdata')
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
```

```{r xvalpt2, message=F, warning=F, echo=T}

performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))

ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

@@

no.hgd.plots = list(); coefs = list(); performance.at.1se = list()
if (file.exists('/tmp/nohgd.Rdata')) {
  load('/tmp/nohgd.Rdata', verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
  for (a in alpha.values) {
    # all patients
    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000)

    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD)

    no.hgd.plots[[as.character(a)]] = arrangeGrob(cv.nohgd$plot+ggtitle('Classification'), cv.nohgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fitNoHGD, cv.nohgd$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nohgd$non.zero.cf)

    performance.at.1se[[as.character(a)]] = subset(cv.nohgd$lambdas, lambda.1se == T)
  }
  save(no.hgd.plots, coefs, performance.at.1se, file='/tmp/nohgd.Rdata')
}

do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r noHGD2, echo=T, message=F, warning=F}
performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))

ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

@@

```{r noLGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
nolgd.plots = list(); coefs = list(); performance.at.1se = list()
if (file.exists('/tmp/nolgd.Rdata')) {
  load('/tmp/nolgd.Rdata', verbose=T)
} else {
  # No LGD
  samples = intersect(rownames(dysplasia.df), c(subset(info, Status == 'NP')$Samplename, subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD') & Status == 'P')$Samplename))

@@

  for (a in alpha.values) {
    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients

    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoLGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoLGD)

    nolgd.plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('Classification'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fitNoLGD, cv.nolgd$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)

    performance.at.1se[[as.character(a)]] = subset(cv.nolgd$lambdas, lambda.1se == T)
  }
  save(nolgd.plots, coefs, performance.at.1se, file='/tmp/nolgd.Rdata')
}

do.call(grid.arrange, c(plots, top="No HGD/IMC/LGD samples", ncol=2))
```

```{r noLGD2, echo=T, message=F, warning=F}
performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))

ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('No HGD/IMC/LGD samples,', folds, 'folds,', splits, 'patient splits'))

@@

```{r, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
m = median(subset(sum.patient.data, Status == 'P')$total.samples)
nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$info$Samplename)))

if (file.exists('/tmp/few.Rdata')) {
  load('/tmp/few.Rdata', verbose=T)
} else {
  # Few
  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples < m)$Patient], function(pt) pt$info$Samplename))))

  few.plots = list()
  for (a in alpha.values) {
    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitF, minR=0.1)
    #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))

    few.plots[[as.character(a)]] = arrangeGrob(cv.back.few$plot+ggtitle('Classification'), cv.back.few$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
  }
  save(few.plots, file='/tmp/few.Rdata')
} 
do.call(grid.arrange, c(few.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples < m))," Progressors with < median # samples (",m,")")))

# Many
if (file.exists('/tmp/many.Rdata')) {
  load('/tmp/many.Rdata', verbose=T)
} else {
  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Patient], function(pt) pt$info$Samplename))))

  many.plots = list()
  for (a in alpha.values) {
    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients

    cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitM$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitM)

    many.plots[[as.character(a)]] = arrangeGrob(cv.back.many$plot+ggtitle('Classification'), cv.back.many$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
  }
  save(many.plots, file='/tmp/many.Rdata')
}

do.call(grid.arrange, c(many.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m))," Progressors with >= median # samples (",m,")")))

@@

info = do.call(rbind, lapply(patient.data, function(df) df$info))
pg.samp = lapply(patient.data, function(pt) {
  info = pt$info
  info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
  info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
  info$Prediction = NA
  info$Prediction.Dev.Resid = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

if (file.exists('/tmp/loo.Rdata')) {
  load('/tmp/loo.Rdata', verbose=T)
} else {
  performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
  # Remove each patient (LOO)
  for (pt in names(pg.samp)) {
    tmp.patient.data = patient.data[subset(sum.patient.data, Patient != pt)$Patient]
    samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))
    length(samples)

    # Fit generated on all samples, including HGD
    a = 1
    fitLOO <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
    cv = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitLOO$lambda, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
    #plots[[pt]] = cv$plot + ggtitle(pt)

    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)

@@

predicted.progression<-function(x, pos=1) { 
  if(length(which(is.na(x))) == length(x)) return(NA)
  as.integer(length(which(x == pos)) > 0)  
}

@@

What about the non-progressors that had progressor predictions?
```{r, warning=F, message=F}

@@

# How do I evaluate the predictions?
